Employee Attrition: What makes an employee leave?

### 2017-10-18 LatexPaper/paper.tex

\paragraph{Several measures were taken to consolidate, sample, and enhance the OPM dataset from its original form. These actions were necessary steps toward prediction readiness. Details are provided in the following subsections.}

@@

\paragraph{Firstly, observations for locations outside the United States were removed as our intent is to model domestic jobs only. In addition, records with no specified occupation, no specified salary, no specified length of service years or level, and no specified age level were removed from the dataset. Age level A employees were also removed as this category includes employees age 20 to as young as age 14, and, therefore, may not reflect relevant work schedules, share common benefits, or separate for valid reasons as compared to the remaining age level categories. Seasonal employees and those not considered full-time were also removed. Finally, all instances with a paygrade (GSEGRD) of less than 7 were dropped. This enabled us to narrow our focus from 150 possible job levels to those that are mid-level and above, and naturally more expensive to train and replace. Note this also removes blue collar workers from our dataset. It should be said that we do not think blue collar positions are less worthy of assessment, but for the purposes of our research, we were most interested in studying positions whose attrition rates would be more costly to an organization.}

\subparagraph{Sufficient difference between occupation types led us to also narrow our selection to only two occupation category types, Professional and Administrative. Fig. 1 depicts the overall record counts present within our dataset by occupation category type. As observed, there are more employees of Administrative and Professional occupation types than any other category. Reviewing separation type ratios which comprise these total counts provides further insight and justification into why Professional and Administrative jobs are targeted within this paper.}

@@

\subparagraph{As Fig. 2[Field] indicates, approximately the same number of Professional employees who retired (SD) in this timeframe quit (SC) as well. Treating retirement as a proxy indicator for retention, this even distribution of quits vs. retirements suggests great retention rates among Professional employees and aligns with non-separation counts as well. Similar argument is to be made of Blue Collar jobs also, and the combination of individual transfers out (SA) and quits is approximately the same in size as retirements for Administrative types. However, the Professional category type was identified as a better candidate for primary model application due to its large total count, as discussed previously, in combination with its more widely diverse professional job types that more closely resemble private sector occupations. Nonetheless, due to its high employee count, strong retention numbers, and similar transfer and quit counts compared to retirements, the Administrative occupation category was selected as secondary target for modelling purposes as will be discussed further in section 5. Therefore, data for all occupation category types, except Professional and Administrative, were removed from our dataset.}

\subparagraph{In addition to observation removal, second and third measures for data preparation include new attribute derivation/acquisition and proper sample design as described in sections 3.2 through 3.4 below.}

>>>>

\paragraph{Our research focuses on the following employee types, only using observations that fit the following rubric: 1) domestic workers (no international positions), 2) employees aged 20 and greater, 3)employees with a job grade level greater than 7 (i.e. white collar only), and only those who are considered full-time employees (i.e. no contracted, temporary, or part-time employees).}

@@

\subparagraph{After reviewing the percentage representation of the various occupation types, as well as investigating the types of jobs that comprise each grouping, we kept only Professional and Administrative positions. Fig. 1 depicts the overall record counts present within our dataset by occupation category type. As observed, there are more employees of Administrative and Professional occupation types than any other category. Reviewing separation type ratios which comprise these total counts provides further insight and justification into why Professional and Administrative jobs are targeted within this paper.}

@@

\subparagraph{As Fig. 2[Field] illustrates, approximately the same number of Professional employees who retired (SD) in this timeframe quit (SC) as well. Treating retirement as a proxy indicator for retention, this even distribution of quits vs. retirements suggests great retention rates among Professional employees and aligns with non-separation counts as well. Similar argument is to be made of Blue Collar jobs also, and the combination of individual transfers out (SA) and quits is approximately the same in size as retirements for Administrative types. However, the Professional category type was identified as a better candidate for primary model application due to its large total count, as discussed previously, in combination with its more widely diverse professional job types that more closely resemble private sector occupations. Nonetheless, due to its high employee count, strong retention numbers, and similar transfer and quit counts compared to retirements, the Administrative occupation category was selected as secondary target for modeling purposes as will be discussed further in section 5. Therefore, data for all occupation category types, except Professional and Administrative, were removed from our dataset.}
